Rust Bug Fix Patterns

### 2022-09-16 introduction.tex
renamed patterns ids, completed descs for 19 patterns, changed to EMSE template, some text for RQ1 is provided

% can we drop the first paragraph altogether?
Producing bugs has been an unavoidable part of computer programming which happens unintentionally. Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software. One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems. Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system program.

@@

% how does this application below differ from the one above?
Similarly, if we are not observing the desired behaviour from our program, that would indicate that there lies a fault within our program. Fault localization is the process of automatically finding the faulty statement(s) in our program. If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize the code segment candidates based on their potentials of producing bugs. That would result in a more robust tool as it leverages from historical data.

Moreover, knowing which code patterns in a programming language are more susceptible to bugs can help programmers to be more cautious while working with similar code structures. This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program. As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.
% maybe more useful for PL design than for programmers; on the other hand I was just talking about stunts in movies and how people are probably more in danger from the less dangerous ones that they do 1000 times rather than the more dangerous ones that they do 2 times; it would be neat if there was a citation on that, but it's a bit of a distraction and doesn't actually help us get our paper in!

% sentence too long; split it.
Moving further afield, researchers have recently been trying to apply deep learning techniques on programs to learn about the semantics of the code or training models which can modify programs towards a specific goal (e.g. bug fix). However, most learning tools get fixed-size vectors as input and map them to outputs. Unlike image data, it is challenging to compress programs into vectors without any semantic information loss. Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms. That results in a large, sparse space for the inputs. Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help prune the input space towards containing fewer dimensions. That can help with creating more efficient models.

% put into a new related work section
% this is why numbers in front of filenames are kind of annoying, eh
Many studies have investigated detecting and categorizing bug fix patterns in general, or for a specific programming language. Our study is focused on pervasive bug fix patterns in the specific language Rust.

% I'll reread this part once it's in its own section.
\paragraph{General bug fix patterns}
Yang et al. proposed a mining approach to detect Python bug fix patterns via studying
fine-grained fixing code changes. In their research, they also examined what portion of the bugs
could be fixed using automated bug fixing approaches. Moreover, they evaluated the fix patterns in
the wild and realized that 37\% of the buggy codes could be matched by the fix patterns they had
found.
Hanam et al. conducted a similar research but for finding Javascript most pervasive bug fix
patterns. They presented a large-scale study of the most common bug fix patterns by mining 105K
commits from 134 server-side JavaScript projects. They used DBSCAN clustering algorithm
and divided bug fixing change types into 219 clusters, from which they extracted 13 pervasive
cross-project bug fix patterns.

\paragraph{Rust bug fix patterns}
In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level
control over resources, as well as high-level safety guarantees. Originally, Rust was a side
project developed by Graydon Hoare, a Mozilla employee. The company started to fund this project
in 2009, since they intended to use Rust in some of the low-level Firefox components.
What distinguishes Rust from the other systems programming languages, e.g. C/C++, is the
ownership framework that has been incorporated inside the language. The ownership framework keeps the track of the references to different memory locations, and will free them once they
go out of scope. If the program passes the compiler checks, it automatically guarantees several
safety problems e.g. dangling references, or double-free problem, that had to be taken care of
manually in languages like C/C++. Moreover, not only the ownership concept removes the
need for a garbage collector in the runtime environment, it does not require the programmer to
explicitly do certain safety tasks, like freeing references.
Due to the memory and thread safety guarantees that Rust provides, many software companies
adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their
storage system, since they were not completely satisfied with the features provided by AWS’s S3 or
Golang.

% ok, now back to the introduction section
% did we talk about ownership yet?
In Rust, the so-called Borrow Checker component enforces ownership rules. Because of the Borrow Checker's importance in Rust, we split patterns into two groups: (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns). %This distinction was really helpful as it made our tool discover clearer results. - can you write something better?

We present our discovered insights of our empirical study in the following four research
questions (RQs):

% we talked about these, right?
\textbf{RQ1. How successful is our data modelling approach in capturing all the necessary data?}

\textbf{RQ2. What are the most common AST structures in the bug fixing changes?}

\textbf{RQ3. What are the general fix patterns in Rust?}

\textbf{RQ4. Which fix patterns are related to Borrow Checker in Rust?}

>>>>

Producing bugs has been an unavoidable part of computer programming. Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software. One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems. Research efforts for categorizing common bug patterns date back to 1975 where Edres tried to categorize various bug patterns which occurred in operating system implementations.

@@

Similarly, if we are observing undesired behaviour from our program, that would indicate that there lies a fault within our program. But where is the fault? Fault localization is the process of automatically finding the faulty statement(s) in our program. If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize code segment candidates based on their potentials of producing bugs. That would result in a more robust tool that leverages historical data.

Moreover, knowing which code patterns in a programming language are more susceptible to bugs can help programmers to be more cautious while working with similar code structures. This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program. As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance.

Moving further afield, researchers have been exploiting deep learning techniques to learn about the semantics of computer programs and create program modifier models that have a specific goal (e.g. bug fix). However, most learning tools get fixed-size vectors as input and map them to outputs. Unlike image data, it is challenging to compress programs into vectors without any semantic information loss. Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms. That results in a large, sparse space for the inputs. Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help prune the input space towards containing fewer dimensions. That can help with creating more efficient models.

Many studies have investigated detecting and categorizing bug fix patterns in general, or for a specific programming language. In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees. What distinguishes Rust from the other systems programming languages, e.g. C/C++, is its ownership framework, which is an integral part of Rust. The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope. If a program passes the compiler’s checks, it must meet several safety properties; e.g. it has no dangling references or double-frees. These problems had to be manually ensured in languages like C/C++. Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks. Due to the memory and thread safety guarantees that Rust provides, many software companies adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang.

In Rust, the so-called Borrow Checker component enforces ownership rules. Because of the Borrow Checker's importance in Rust, we split patterns into two groups: (1) non-borrow-checker related patterns (General Patterns), and (2) borrow-checker related patterns (BC-Related Patterns). We present our discovered insights of our empirical study in the following four research
questions (RQs):

\textbf{RQ1. How successful is our data modelling approach in representing program change?}

\textbf{RQ2. What are the general fix patterns in Rust, and how often do they apply?}

\textbf{RQ3. What are borrow-checker related fix patterns in Rust, and how often do they apply?}

### 2022-09-16 results.tex
renamed patterns ids, completed descs for 19 patterns, changed to EMSE template, some text for RQ1 is provided

% some general comments about the taxonomy go here

% take one example and explain it, e.g. ++ means ..., -- means ....
% rename to e.g. B.ref.add
% need quantitative data

@@

Attributes in Rust make it possible for the developers to add metadata to program elements. These attributes can be added to structs, struct fields, enum variants, match expression arms, etc. There are two types of attributes: Inner Attributes and Outer Attributes. Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute. % does the inner/outer distinction matter?

\subsubsection{G1. Modifying the attributes of structs}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G1.1: Adding attributes to a struct
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

// G1.2: Removing attributes from a struct
-- #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

// G1.3: Changing an existing attribute of a struct
-- #[derive(Debug)]
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}
\end{lstlisting}

\noindent \textbf{Description:} In this pattern, the attribute set of a struct changes. This modification can happen in three different ways; (G1.1) Adding attributes to the struct, (G1.2) Removing attributes from the struct, (G1.3) Changing the content of an existing attribute. The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern.

\subsubsection{G2. Modifying the attributes of struct fields}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G2.1: Adding attributes to struct fields
#[derive(Serialize)]
pub struct Retain {
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}

// G2.2: Removing attributes from struct fields
#[derive(Serialize)]
pub struct Retain {
    -- #[serde(skip_serializing_if = "Option::is_none")]
    pub attributes: Option<Attributes>,
}

// G2.3: Changing attributes of struct fields
#[derive(Serialize)]
pub struct Retain {
    -- #[serde(skip_serializing_if = "Option::is_none")]
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}
\end{lstlisting}

% The sentence "Using..." doesn't actually have an appropriate verb
\noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes. As with G1, this modification also happens in three different ways. Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.

\subsection{Struct}
\subsubsection{G3. Adding/Removing a Struct field}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G3.1: Adding a new field to a struct
struct InnerListeners {
    pending: Mutex<Vec<Pending>>,
    <@\textcolor{red}{--}@> queue_object_name: Uuid,
}

// G3.2: Removing a field from a struct
struct SnapshotService<U, R> {
    uuid_resolver_handle: R,
    <@\textcolor{green}{++}@> db_name: String,
}
\end{lstlisting}

\noindent \textbf{Description:} A set of new fields are add to or removed from an existing struct. The addition of a new field happens when the developer need to store a new piece of data so they can implement a desired behaviour. The removal of a field  happens when the developer realizes that the field is expendable with respect to the behaviour they want to implement.

\subsubsection{G4. Modifying the access modifiers of struct fields}

@@

// G4.1: Adding pub to a field
pub struct HtmlBlock {
    -- content: BlockContent,
    ++ pub content: BlockContent,
    brace: token::Brace,
}

// G4.2: Removing pub from a field
pub struct Languages {
--  pub named: HashMap<LanguageId, Arc<LanguageDefinition>>,
++  named: HashMap<LanguageId, Arc<LanguageDefinition>>,
    extensions: HashMap<String, Arc<LanguageDefinition>>,
}

@@

\subsubsection{G5. Different fields order when initializing Struct}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- return Poll::Ready(ReadyEvent { ready, tick });
++ return Poll::Ready(ReadyEvent { tick, ready });

\noindent \textbf{Description:} Remove... ???

\subsection{Option}
\subsubsection{G6. Changing field type T to T$\langle$Option$\rangle$}

@@

\subsection{Types}
\subsubsection{G7. Changing a struct field type}

@@

\noindent \textbf{Description:} The type of a struct field changes. This is a common change pattern in all statically-typed languages and can happen for various reasons. These reasons can be a step for implementing a new feature or rectifying the behaviour of the program. Also it can happen for refactoring purposes or performance enhancement.

\subsubsection{G8. Changing a generic type}

#[derive(Debug, Clone, PartialEq)]
pub struct Anchor {
    -- point: Point<usize>,
    ++ point: Point<isize>,
    side: Side,
}

@@

\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to G7, this change can also happen for the mentioned reasons.

\subsubsection{G9. Change in enum variant value type}

#[derive(Debug, Clone, PartialEq, Eq)]
pub enum ResolvedDependency {
  Resolved(ModuleSpecifier),
  -- Err(String),
  ++ Err(ResolvedDependencyErr),
}

@@

\noindent \textbf{Description:} In Rust, we can specify the type of the value that an enum variant can store. Using this feature, we can avoid using a struct alongside our enum variants to specify the value type. This change pattern is associated with a change in enum variant types.

\subsubsection{G10. Changing the type inside a tuple}

#[allow(dead_code)]
pub struct CoreState {
    // Other fields
    -- pending_views: Vec<ViewId>,
    ++ pending_views: Vec<(ViewId, Table)>,
    peer: Client
}

@@

\subsection{Traits}
\subsubsection{G11. TraitBound Change}

-- pub struct EventLoop<T: tty::EventedReadWrite> {
++ pub struct EventLoop<T: tty::EventedPty> {
    poll: mio::Poll,
    // Other fields
}

@@

\subsubsection{G12. Implementing a different trait for a struct}

-- impl ToDebug for ShellError {
++ impl FormatDebug for ShellError {
        // Implementing functions ...
   }

\noindent \textbf{Description:} Remove... ???

\subsection{Ref and Deref}
\subsubsection{B1. Adding $\&$}

@@

To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}, file \texttt{git\_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope. Now this object can then be used in another function invocation ($remove\_dir\_all(repo\_dir)$) which fixes a bug.

\subsubsection{B2. Removing $\&$}

@@

\noindent \textbf{Description:} In the pattern, an unnecessary borrowing is removed. Clippy can detect the needless borrows, and also can automatically remove them. This does not change the behaviour of the program, however, it does improve the performance, as the borrow checker need not to return the ownership after the execution of the function.

\subsubsection{B3. Dropping \& from func decl and adding * to match statement}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- fn to_color_choice(&self) -> ColorChoice {
--     match *self {
++ fn to_color_choice(self) -> ColorChoice {
++     match self {
           // Statements
       }
\end{lstlisting}

\noindent \textbf{Description:} ???

\subsubsection{B4. Modification in dereferencing a string prior to calling .to\_string()}

@@

\subsection{Mut}
\subsubsection{B5. Adding mut}

parser::Parser::new(args)
        .parse_module()
--      .map_err(|e| {
++      .map_err(|mut e| {
            e.emit();
            ()
        })?

@@

\subsubsection{B6. Removing mut}

@@

\subsection{Vector}
\subsubsection{B7. Vector reference to slice}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
pub fn expand_delimited_square(
--  children: &Vec<TokenNode>,
++  children: &[TokenNode],
) -> Result<hir::Expression, ParseError> {
    // body
}
\end{lstlisting}

\noindent \textbf{Description:} ???

\subsection{Lifetime}
\subsubsection{B8. Adding generic Lifetime}
\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
#[allow(dead_code)]
-- fn api_error(webview: &mut Webview, error_fn: String, message: &str) {
++ fn api_error(webview: &mut Webview<'_>, error_fn: String, message: &str) {
  let reject_code = tauri_api::rpc::format_callback(error_fn, message);
  webview.eval(&reject_code)
}

@@

\noindent \textbf{Description:} ???

\subsubsection{B9. Dropping static lifetime}

-- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];

\noindent \textbf{Description:} ???

\subsubsection{B10. Dropping Lifetime in field definition}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
pub struct Table<'a> {
    columns: Vec<Column>,
    colours: &'a Colours,
    env: &'a Environment,
    widths: TableWidths,
    -- time_format: &'a TimeFormat,
    ++ time_format: TimeFormat,
    size_format: SizeFormat,
    git: Option<&'a GitCache>,
}
\end{lstlisting}

\noindent \textbf{Description:} ???

\subsection{Clone}
\subsubsection{B11. Dropping clone (and adding dereference)}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
pub fn reset(&mut self, template: &Cell) {
    // memcpy template to self
--  *self = template.clone();
++  *self = *template;
}
\end{lstlisting}

\subsubsection{B12. Dropping clone}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
if let Some(var_id) = &var.var_id {
--  stack.add_var(*var_id, input);
++  stack.add_var(*var_id, input.clone());
}

start_plugin_process(
--  manifest.clone(),
++  manifest,
    self.next_plugin_id(),
    self.self_ref.as_ref().unwrap().clone(),
);

\subsubsection{B13. Dropping clone and adding reference}

Similar to the previous pattern, this does not change the program behaviour and is merely for performance purposes. It was the case in one of the commits of the project tauri-apps/tauri, where the clone to borrowing significantly optimized CPU usage. They discussed that the cloning of some types such as HashMap was expensive.

@@

\hskip-2.5cm\begin{tabular}{|l|l|l|l|}
\hline

\multirow{12}{*}{General}        
& \multirow{2}{*}{Attributes}    & G1  & Adding Attributes to Structs                                        \\
                                 &                                & G2  & Adding Attributes to Struct Fields                                  \\\cline{2-3}
                                 & \multirow{3}{*}{Struct}        & G3  & Adding/Removing a Struct field                                      \\
                                 &                                & G4  & Adding/Removing pub to/from struct field                            \\
                                 &                                & G5  & Different fields order when initializinng Struct                       \\\cline{2-3}
                                 & Option                         & G6  & Changing field type T to Option                                          \\\cline{2-3}
                                 & \multirow{4}{*}{Types}         & G7  & Changing a Struct field type                                        \\
                                 &                                & G8  & Changing a generic type                                             \\
                                 &                                & G9  & Change in enum variant value type                                   \\
                                 &                                & G10 & Changing the type inside a tuple                                         \\\cline{2-3}
                                 & \multirow{2}{*}{Traits}        & G11 & TraitBound Change                                                   \\
                                 &                                & G12 & Implementing a different trait for a struct.                            \\\cline{1-3}
\multirow{13}{*}{Borrow Checker} & \multirow{4}{*}{Ref and Deref} & B1  & Adding \&                                                           \\
                                 &                                & B2  & Removing \&                                                         \\
                                 &                                & B3  & Dropping \& from function declaration and dereferencing match \\
                                 &                                & B4  & Adding *                                                            \\\cline{2-3}
                                 & \multirow{2}{*}{Mut}           & B5  & Adding mut                                                          \\
                                 &                                & B6  & Removing mut                                                        \\\cline{2-3}
                                 & Vector                         & B7  & Vector reference to slice                                           \\\cline{2-3}
                                 & \multirow{3}{*}{Lifetime}      & B8  & Adding generic Lifetime                                             \\
                                 &                                & B9  & Dropping static lifetime                                            \\
                                 &                                & B10 & Dropping Lifetime in field definition                               \\\cline{2-3}
                                 & \multirow{3}{*}{Clone}         & B11 & Dropping clone and adding dereference                               \\
                                 &                                & B12 & Dropping clone                                                      \\
                                 &                                & B13 & Dropping clone and adding reference  \&

>>>>

\subsection{Evaluating the data modelling approach (RQ1)}

As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fix-sized datapoints which then could be used in the DBSCAN clustering algorithm. We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST. Here we want to evaluate the effectiveness of our encoding.

The general idea for our evaluation is whether the most important program elements has the high values in our encoding. The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point. We used circle pack figures to visualize each data point. In these figures, the radius of the circle of each non terminal shows its importance in our encoding (Fig ref).

We choose 50 random samples from each of our general and borrow-checker databases. These random samples are drawn from the clustered datapoints. The reason behind this decision is the same as why we collected a lot of noise points, that is, the majority of the commits contain multi-hunk complex changes and the key changing elements are difficult to determine.

We carried out two manual experiments for evaluating our encoding mechanism. In the first experiment, the analyzer investigates whether the most important element has the largest radius or not. In the second experiment, the analyzer investigates whether the most important element is included in three largest circles or not. The analyzer should give a binary answer to these questions. Both authors carry out these experiments independently. At last, we report the results as the portion of the changes which both authors agreed upon its encoding correctness. The final results are TBD\% and TBD\% correctness in the first and second experiments, respectively, which shows the effectiveness of our encoding mechanism.

\subsection{Common Bug Fix Patterns (RQ2)}

@@

Attributes in Rust make it possible for the developers to add metadata to program elements. These attributes can be added to structs, struct fields, enum variants, match expression arms, etc. There are two types of attributes: Inner Attributes and Outer Attributes. Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute. \\

\noindent \textit{ \textbf{G.attr.struct.} Modifying the attributes of structs}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G.attr.struct.add: Adding attributes to a struct
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

//G.attr.struct.drop: Removing attributes from a struct
-- #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

// G.attr.struct.change: Changing an existing attribute of a struct
-- #[derive(Debug)]
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}
\end{lstlisting}

\noindent \textbf{Description:} In this pattern, the attribute set of a struct changes. This modification can happen in three different ways; (G1.1) Adding attributes to the struct, (G1.2) Removing attributes from the struct, (G1.3) Changing the content of an existing attribute. The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern. \\

\noindent \textit{ \textbf{G.attr.struct.field} Modifying the attributes of struct fields}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G.attr.struct.field.add: Adding attributes to struct fields
#[derive(Serialize)]
pub struct Retain {
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}

// G.attr.struct.field.drop: Removing attributes from struct fields
#[derive(Serialize)]
pub struct Retain {
    -- #[serde(skip_serializing_if = "Option::is_none")]
    pub attributes: Option<Attributes>,
}

// G.attr.struct.field.change: Changing attributes of struct fields
#[derive(Serialize)]
pub struct Retain {
    -- #[serde(skip_serializing_if = "Option::is_none")]
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}
\end{lstlisting}

\noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes. As with G1, this modification also happens in three different ways. Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.

\subsection{Struct} 
\noindent \textit{ \textbf{G.struct.field} Adding/Removing a struct field}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G.struct.field.add: Adding a new field to a struct
struct InnerListeners {
    pending: Mutex<Vec<Pending>>,
    <@\textcolor{red}{--}@> queue_object_name: Uuid,
}

// G.struct.field.drop: Removing a field from a struct
struct SnapshotService<U, R> {
    uuid_resolver_handle: R,
    <@\textcolor{green}{++}@> db_name: String,
}
\end{lstlisting}

\noindent \textbf{Description:} A set of new fields are add to or removed from an existing struct. The addition of a new field happens when the developer need to store a new piece of data so they can implement a desired behaviour. The removal of a field  happens when the developer realizes that the field is expendable with respect to the behaviour they want to implement. \\

\noindent \textit{ \textbf{G.struct.field.pub} Modifying the access modifiers of struct fields}

@@

// G.struct.field.pub.add: Adding pub to a field
pub struct HtmlBlock {
    -- content: BlockContent,
    ++ pub content: BlockContent,
    brace: token::Brace,
}

// G.struct.field.pub.drop: Removing pub from a field
pub struct Languages {
--  pub named: HashMap<LanguageId, Arc<LanguageDefinition>>,
++  named: HashMap<LanguageId, Arc<LanguageDefinition>>,
    extensions: HashMap<String, Arc<LanguageDefinition>>,
}

@@

\subsection{Option}
\vspace{3mm}

\noindent \textit{ \textbf{G.option} Changing field type T to T$\langle$Option$\rangle$}

@@

\subsection{Types}
\vspace{3mm}

\noindent \textit{ \textbf{G.type.field} Changing a struct field type}

@@

\noindent \textbf{Description:} The type of a struct field changes. This is a common change pattern in all statically-typed languages and can happen for various reasons. These reasons can be a step for implementing a new feature or rectifying the behaviour of the program. Also it can happen for refactoring purposes or performance enhancement. \\

\noindent \textit{ \textbf{G.type.generic} Changing a generic type}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
#[derive(Debug, Clone, PartialEq)]
pub struct Anchor {
    -- point: Point<usize>,
    ++ point: Point<isize>,
    side: Side,
}

@@

\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to G7, this change can also happen for the mentioned reasons. \\

\noindent \textit{ \textbf{G.type.enum.variant} Change in enum variant value type}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
#[derive(Debug, Clone, PartialEq, Eq)]
pub enum ResolvedDependency {
  Resolved(ModuleSpecifier),
  -- Err(String),
  ++ Err(ResolvedDependencyErr),
}

@@

\noindent \textbf{Description:} In Rust, we can specify the type of the value that an enum variant can store. Using this feature, we can avoid using a struct alongside our enum variants to specify the value type. This change pattern is associated with a change in enum variant types. \\

\noindent \textit{ \textbf{G.type.tuple} Changing the type inside a tuple}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
#[allow(dead_code)]
pub struct CoreState {
    // Other fields
    -- pending_views: Vec<ViewId>,
    ++ pending_views: Vec<(ViewId, Table)>,
    peer: Client
}
\end{lstlisting}

@@

\subsection{Traits}
\noindent \textit{ \textbf{G.trait.bound} TraitBound Change}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- pub struct EventLoop<T: tty::EventedReadWrite> {
++ pub struct EventLoop<T: tty::EventedPty> {
    poll: mio::Poll,
    // Other fields
}

@@

\subsection{BC-Related Bug Fix Patterns (RQ3)}

\subsubsection{Clone}

\noindent \textit{ \textbf{BC.clone.ref} Dropping clone and adding borrowing}

Similar to the previous pattern, this does not change the program behaviour and is merely for performance purposes. It was the case in one of the commits of the project tauri-apps/tauri, where the clone to borrowing significantly optimized CPU usage. They discussed that the cloning of some types such as HashMap was expensive. \\

\noindent \textit{ \textbf{BC.clone.drop} Removing clone}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
start_plugin_process(
--  manifest.clone(),
++  manifest,
    self.next_plugin_id(),
    self.self_ref.as_ref().unwrap().clone(),
);
\end{lstlisting}

\subsubsection{Ref and Deref}

\noindent \textit{ \textbf{BC.ref.add} Adding Borrowing}

@@

To give an example, in starship/starship after one of the commits \footnote{\url{https://github.com/starship/starship/commit/56d475578ea508631275772127f49a6949fea6b0}{git_branches.rs}}, the developer borrows the value of $repo\_dir$ to keep the ownership of the object in the current scope. Now this object can then be used in another function invocation ($remove\_dir\_all(repo\_dir)$) which fixes a bug. \\

\noindent \textit{ \textbf{BC.ref.drop} Removing Borrowing}

@@

\noindent \textbf{Description:} In the pattern, an unnecessary borrowing is removed. Clippy can detect the needless borrows, and also can automatically remove them. This does not change the behaviour of the program, however, it does improve the performance, as the borrow checker need not to return the ownership after the execution of the function. \\

\noindent \textit{ \textbf{BC.deref.to\_string()} Modification in dereferencing a string prior to calling to\_string()}

@@

\subsubsection{Mut}

\noindent \textit{ \textbf{BC.mut.add} Adding mutability}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
parser::Parser::new(args)
        .parse_module()
--      .map_err(|e| {
++      .map_err(|mut e| {
            e.emit();
            ()
        })?
\end{lstlisting}

@@

\noindent \textit{ \textbf{BC.mut.drop} Dropping mutability}

@@

\subsubsection{Vector}

\noindent \textit{ \textbf{BC.vec.slice} Changing a Vec reference to Slice}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
pub fn expand_delimited_square(
--  children: &Vec<TokenNode>,
++  children: &[TokenNode],
) -> Result<hir::Expression, ParseError> {
    // body
}
\end{lstlisting}
\noindent \textbf{Description:} In struct field definition or the types of the function formal parameters, type $\&Vec\<T\>$ changes to $\&[T]$. Clippy can detect the opportunity to apply this pattern. The reason is reducing type complication, as slice types in the form of $\&[T]$ or $\&str$ are usually sufficient for most use cases.

\subsubsection{Lifetime}
\noindent \textit{ \textbf{BC.lifetime.static.drop} Dropping static lifetime}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];

@@

\noindent \textbf{Description:} Removing the static keyword from a const variable declaration. This change can be applied using Clippy. If the presence of static is not required, it is better that we omit it as keeping it might create complicate types in our program.

@@

\begin{tabular}{|l|l|l|l|}
\hline

\multirow{10}{*}{General}        
& \multirow{2}{*}{Attributes}    & G.attr.struct  & Modifying the attributes of structs                                        \\
                                 &                                & G.attr.struct.field  & Modifying the attributes of struct fields                                  \\\cline{2-3}
                                 & \multirow{2}{*}{Struct}        & G.struct.field  & Adding/Removing a struct field                                      \\
                                 &                                & G.struct.field.pub  & Modifying the access modifiers of struct fields     
                                  \\\cline{2-3}
                                 & Option                         & G.option  & Changing field type T to T$\langle$Option$\rangle$                     \\\cline{2-3}
                                 & \multirow{4}{*}{Types}         & G.type.field  & Changing a struct field type                                        \\
                                 &                                & G.type.generic  & Changing a generic type                                             \\
                                 &                                & G.type.enum.variant  & Change in enum variant value type                                   \\
                                 &                                & G.type.tuple & Changing the type inside a tuple                                         \\\cline{2-3}
                                 & \multirow{1}{*}{Traits}        & G.trait.bound & TraitBound Change                                               \\\cline{1-3}
\multirow{9}{*}{Borrow Checker} &
\multirow{2}{*}{Clone}                                         & BC.clone.ref & Dropping clone and adding borrowing                                                   \\
                                 &                                & BC.clone.drop & Dropping clone   
                                    \\\cline{2-3}
    & \multirow{3}{*}{Ref and Deref} & BC.ref.add  & Adding Borrowing                                                           \\
                                 &                                & BC.ref.drop  & Removing Borrowing                                                         \\
                                 &                                & BC.deref.to\_string  & Modification in dereferencing a string prior to calling to\_string()                                                            \\\cline{2-3}
                                 & \multirow{2}{*}{Mut}           & BC.mut.add  & Adding mutability                                                          \\
                                 &                                & BC.mut.drop  & Dropping mutability                                                        \\\cline{2-3}
                                 & Vector                         & BC.vec.slice  & Changing a Vec reference to Slice                                           \\\cline{2-3}
                                 & \multirow{1}{*}{Lifetime}      
                                                                 & BC.lifetime.static  & Dropping static lifetime    &                        
                                 \hline
